---
title: "LotekFilter-Advanced"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LotekFilter-Advanced}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 7,
  fig.width = 7)
```
This tutorial will cover the details of the filter criteria and alternative ways to utilize the package functions. See 'Recommended Usage' vignette for basic usage instructions. The LotekFilter package was created based on TXT files exported by WHS Host V1.9.4817.2 (March 17, 2023) using Tools -> 'Batch data to text conversion'. 

While the LotekFilter R package was designed to parallelize the filtering process across multiple input text files, it is possible to execute the parallel_ functions on a single file. If you are multi-tasking on the computer that will be processing the data, it is better to reserve more unused cores (e.g. n_cores = parallel::detectCores()/2). If you want to maximize processing speed and can minimize other tasks and programs executed on the computer while performing the filtering, more cores can be committed (e.g. n_cores = parallel::detectCores()-1). Always reserve at least one core unassigned for basic operating system functions. 

## Install
To install `LotekFilter` package directly from GitHub:
```{r install, echo=TRUE, eval=TRUE}
devtools::install_github("dalcottcdfw/LotekFilter") # install from GitHub
```

## Filtering Process
Filter steps:
1. Allowable tag codes only (optional)
2. Remove multipath detections (i.e. multiple records of the same code within fractions of a second of one another) 
3. Pulse Rate Interval filter (per tag code)
  a. at least 'min_detections' detections within 'detection_window'-fold of the nominal PRI in seconds (default: 4 detections within 16.6-fold of nominal PRI)
  b. difference between observed pulse rate and nominal pulse rate muse be <'nominalPRI_threshold' (default: 0.20 or 20%)
  c. standard deviation of calculated pulse rate from 3 consecutive detections must be <'sd_threshold' (default: 0.025)
  
The first filtering step is filtering to only allow known tag codes. This is not strictly required if it is possible that unknown tag codes may be detected. However, filtering to only allow known tag codes has the largest reduction on file size and takes place at the beginning of the pipeline, resulting in dramatic performance improvements. If your tag list is in decimal format instead of the required hexidecimal format, the dec2hex() function from the broman package can convert decimal format codes to hexidecimal. The converted codes will need to be forced to uppercase for proper matching.
```{r code conversion, eval = FALSE}
HexCodes <- toupper(broman::dec2hex(DecimalCodes)) # example conversion of decimal codes to hexidecimal.
```


The next step is to remove multipath detections. These are detections of the same tag code within a very short amount of time of each other. By default, 0.3 seconds is used as the threshold. All records of a tag within 0.3 seconds of the same tag on the same receiver will be removed as a multipath detection. This prevents multipath detections from interfering with the PRI filter assessment. 

The default PRI filter criteria are derived from criteria used by Arnold Ammann for use in California's Central Valley JSATS telemetry database hosted on NOAA's ERDDAP server. These filter criteria have been tuned for Lotek receivers specifically and perform fairly well at removing false-positive records without removing too many true-positives. Some false-positives are expected after filtering. Additional filter criteria such as reasonable movement constraints, expected tag life expectancy, and known receiver deployment periods will likely be necessary. A given PRI filter criterion could be effectively removed by making the parameter extremely tolerant. For example, if you do not want to use the standard deviation of calculated pulse rate criterion, 'sd_threshold' can be set to a very large value (e.g. 999999). Reducing min_detections to <3 will likely result in a very large number of false-positives retained in the final data after filtering. 

## Altering default criteria
The example below alters several of the default filter criteria and applies a single nominal PRI to all tags. These criteria have not been tested and are shown for example purposes only. 
```{r new criteria}
library(LotekFilter)
input_files <- list.files(pattern = "Processed_*")
output_path <- "C:/Documents/jsats_study/data/"
result_summary <- parallel_filter_Lotek <- function(input_files,
                                  input_prefix = "Processed_", 
                                  output_prefix = "Filtered_",
                                  output_path = output_path,
                                  
                                  keep_rejected = TRUE, # TRUE if you want to save a copy of the rejected records with a reason for rejecting
                                  n_cores    = min(length(input_files), 
                                                   parallel::detectCores()-2), # 2 cores less than all available or number of files, whatever is smaller
                                  nominalPRI = 5,           # provide a single numeric tag pulse rate interval to all tags
                                  detection_window = 19,    # detection_window * nominmal PRI = time window that min_detections must occur to be kept
                                  min_detections = 5,         # number of detections required within the detection window to be kept.
                                  sd_threshold = 99999,           # max threshold for standard deviation of observed PRI
                                  multipath_threshold = 0.5,  # time threshold for multipath detections. Detections less than this after first are removed. 
                                  nominalPRI_threshold = 0.2 # observed PRI must be within this amount of nominal PRI.
)
```


## Filtering non-Lotek files
While the parallel_raw_Lotek() function was designed specifically for processing the raw text output files from Lotek WHS Host and receivers, the parallel_filter_Lotek() function could be applied to detection records from any manufacturer, as long as the data are configured properly. In order to filter out false-positive detections from data files that do not match the formatting from Lotek's WHS Host, the user would need to create one csv file for each receiver that contains the following columns:
'HexID' - column containing the 4 digit hexidecimal tag code
'DateTime' - column containing the datetime of the detection in yyyy-mm-dd hh:mm:ss format.
The standard filtering criteria for other receiver models (e.g. ATS, Teknologic) differ from the Lotek filtering criteria. Most of these standard filtering criteria can be captured using the filter settings arguments in the [parallel_filter()] function. However, the [parallel_filter()] function does not capture the frequency filter criterion commonly applied to ATS and Teknologic receivers. That filter would have to be applied manually afterwards.
```{r Filter Non-Lotek}
library(LotekFilter)
# Hypothetical example filtering csv files derived from ATS receivers (altered to contain necessary columns)

input_files <- list.files(pattern = "ATS_inputs_*")
output_path <- "C:/Documents/jsats_study/data/"

result_summary <- parallel_filter_Lotek <- function(input_files,
                                  input_prefix = "Processed_", 
                                  output_prefix = "Filtered_",
                                  output_path = output_path,
                                  
                                  keep_rejected = TRUE, # TRUE if you want to save a copy of the rejected records with a reason for rejecting
                                  n_cores    = min(length(input_files), 
                                                   parallel::detectCores()-2), # 2 cores less than all available or number of files, whatever is smaller
                                  nominalPRI = 5,           # provide a single numeric tag pulse rate interval to all tags
                                  
                                  # Standard ATS filter criteria:
                                  detection_window = 4,    # detection_window * nominmal PRI = time window that min_detections must occur to be kept
                                  min_detections = 2,         # standard ATS minimum detections
                                  sd_threshold = 99999,           # ATS criteria do not have sd threshold - use permissively large threshold to effectively remove this criterion
                                  multipath_threshold = 0.3,  # time threshold for multipath detections. Detections less than this after first are removed. 
                                  nominalPRI_threshold = 0.1 # standard ATS PRI tolerance
)

## Additional ATS frequency filter criteria:
  # frequency must be between 416.3 - 418.75 kHz
  # difference in frequency between 2 hits is less than 0.505 kHz
```



## Alternate use cases: Single file
The LotekFilter package was designed to process batches of raw Lotek TXT files efficienctly. It is possible to use the parallel functions with only a single file. This will attempt to use furrr and future to map parallel processing, but only one core will be assigned. However, it is also possible to bypass the parallel functions and apply the single file functions to a single file. In order to run a single file this way, you must specify all filter criteria as arguments in the function call. This is an example of processing then filtering a single file:
```{r single file}
library(LotekFilter)
tags <- read.csv(paste(file.path(system.file("extdata", package = "LotekFilter"), "example_tagcodes.csv"), sep = "/"))# read in example tag codes 
# unzip the zipped detections data
zip_path <- system.file("extdata", "ExampleDetectionsFiles.zip", package = "LotekFilter")
tmpdir <- tempdir()
unzip(zip_path, exdir = tmpdir)
list.files(tmpdir, pattern = ".TXT") # check that the detection test files are in the zip folder, should contain two "WHS4k..."files
output_dir = system.file("extdata", package = "LotekFilter")


# Convert a single raw Lotek file to csv
process_single_raw(raw_file = list.files(tmpdir, pattern = ".TXT")[1], # first example file from package extdata
                               input_path = tmpdir, # temporary directory for zip file contents (change as needed)
                               output_path = output_dir,
                               output_prefix = "",
                               allowable_tagcodes = tags$TagID_Hex,
                               tz = "Etc/GMT+8")


# False-positive filter a single Lotek file
settings = list(inputpath = output_dir,
                output_path = output_dir,
                input_prefix = "WHS4k-",
                output_prefix = "Filtered_",
                
                # if running single file: must specify all filter criteria
                nominalPRI = 5,
                detection_window = 16.6,    
                min_detections = 4,
                sd_threshold = 0.025,
                multipath_threshold = 0.3,
                nominalPRI_threshold = 0.2,
                
                keep_rejected = FALSE)

input_file <- paste(output_dir, sub("\\.TXT$", ".csv", list.files(tmpdir, pattern = ".TXT")[1]), sep = "/")
process_single_file(filepath = input_file, 
                    settings = settings) 
#quietly exports a filtered csv to the output directory

```

