---
title: "LotekFilter-Recommended"
author: "Derrick Alcott"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{LotekFilter-Recommended}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
echo = FALSE,
message = FALSE,
warning = FALSE,
fig.height = 7,
fig.width = 7)
```
This tutorial will describe the recommended and most common usage of the `LotekFilter` package. See 'Advanced Usage' vignette for all package functionality. 

## Package purpose
Lotek JSATS receivers generate large, encrypted .JST files. The encrypted .JST files can be converted to accessible .TXT files using Lotek's WHS Host software. The LotekFilter package was created based on TXT files exported by WHS Host V1.9.4817.2 (March 17, 2023) using Tools -> 'Batch data to text conversion'. Those text files are not computation-friendly as they contain multiple different data tables in the same file and each data table is fixed width format. Furthermore, JSATS detection files are often riddled with many false-positive detections, which can lead to very large data files over the course of a single receiver deployment (often >1 million records per receiver). The purpose of the `LotekFilter` package is to efficiently convert the data files produced by WHS Host software into a nearly ready-for-analysis, user-friendly condition. Note: JSATS detections will always need careful quality control checks before inference. 

## Install
To install `LotekFilter` package directly from GitHub:
```{r install, echo=TRUE, eval=TRUE}
devtools::install_github("dalcottcdfw/LotekFilter") # install from GitHub
```

## Example workflow
Here's a quick example workflow of how the package was designed to function in most cases:
```{r workflow, echo=TRUE, eval=FALSE}
library(LotekFilter)
input_dir <- "C:/Documents/jsats_study/data_raw/" # location where input files are stored
my_files <- list.files(input_dir, pattern = "\\.TXT$") # choose all text files in the input directory as the input files
output_dir<- "C:/Documents/jsats_study/data/" # location to save the processed output files
allowable_tags <- read.csv("C:/Documents/jsats_study/data/my_allowable_tag_codes.csv") # read in a list of allowable tag codes with their PRI

# Step 1 - Process raw files
parallel_raw_Lotek(input_path = input_dir,
                   raw_files = list.files(input_dir, pattern = "\\.TXT$"), # assumes all .TXT files in input_path are to be processed
                   output_path = output_dir,
                   allowable_tagcodes = allowable_tags$TagID_Hex, # optional but recommended: pre-filter on set of allowable tag codes
                   tz = "Etc/GMT+8", # time zone that detections were recorded in
                   n_cores = parallel::detectCores()-2) # number of computer cores to use (always reserve at least 1 unused core)


## Step 2 - Filter false-positives:
setwd("C:/Documents/jsats_study/data/") # set the working directory to the location where the pre-processed csv files are stored
parallel_filter_Lotek(
  input_files = list.files(pattern = ".csv"),
  output_prefix = "Filtered_",
  output_path = "C:/Documents/jsats_study/data/", # store filtered files in the same location. inputs will not be overwritten because of file prefix
  keep_rejected = FALSE, # TRUE if you want to save a copy of the rejected records with a reason for rejecting
  n_cores    = n_cores = parallel::detectCores()-2, # number of cores to use for parallel processing (1 input file per core, do not assign all cores)
  
  pri_table = allowable_tags,            # provide name of tag pri lookup table AND
  pri_tag_col = "TagID_Hex",          # provide name of hex tag code column in tag lookup table AND
  pri_value_col = "PRI_nominal",        # provide name of column containing tag pulse rate interval in tags lookup table
  
)
```

## Functional example using provided data
This is a functional example that uses example data files provided with the package download. 
```{r echo=TRUE, eval=FALSE}
library(LotekFilter)
output_dir <- "C:/Desktop/path/to/store/outputs" # change to your preferred output storage location

tags <- read.csv(paste(file.path(system.file("extdata", package = "LotekFilter"), "example_tagcodes.csv"), sep = "/"))# read in example tag codes data

# unzip the zipped detections data
zip_path <- system.file("extdata", "ExampleDetectionsFiles.zip", package = "LotekFilter")
tmpdir <- tempdir()
unzip(zip_path, exdir = tmpdir)
list.files(tmpdir, pattern = ".TXT") # check that the detection test files are in the zip folder, should contain two "WHS4k..."files

# Step 1 - process raw files
parallel_raw_Lotek(input_path = tmpdir,
                   raw_files = list.files(tmpdir, pattern = "\\.TXT$"),
                   output_path = output_dir,
                   output_prefix = "Processed_",
                   allowable_tagcodes = tags$TagID_Hex,
                   tz = "Etc/GMT+8",
                   n_cores = 2)


# Step 2 - filter out false-positive detections
setwd(output_dir) # set directory to location where processed files are stored
parallel_filter_Lotek(
  input_files = list.files(pattern = ".csv"),
  input_prefix = "Processed_", # if the input files have a prefix that you want to remove when saving the output ("Raw_..." or "Prefiltered_...")
  output_prefix = "Filtered_",
  output_path = output_dir,
  
  n_cores    = 2, # number of cores to use for parallel processing (1 input file per core)
  
  pri_table = tags,            # provide name of tag pri lookup table AND
  pri_tag_col = "TagID_Hex",          # provide name of hex tag code column in tag lookup table AND
  pri_value_col = "PRI_nominal",        # provide name of column containing tag pulse rate interval in tags lookup table
)

# Step 3 - combine all filtered files into one dataframe object in R environment
filtered_files <- list.files(output_dir, pattern = "Filtered_*")


# Read each file and add a filename column
df_list <- lapply(filtered_files, function(file) {
  df <- read.csv(file)
  df$Filename <- basename(file) # add source file name for identifying receiver
  df
})

# Combine into one data frame
df <- do.call(rbind, df_list)
# df is a final dataframe object with all filtered detections

```
It is highly recommended to use a vector of allowable tag codes to filter the raw detections data as early as possible. This will greatly speed up the rest of the process. However, it is possible to skip this step.
These examples use default filter criteria which are based on the filter parameters applied by Arnold Ammann for use in the California Central Valley JSATS telemetry database hosted on NOAA's ERDDAP server (see [Ammann 2020; Environ Biol Fish](https://www.bing.com/ck/a?!&&p=4baefa46e97731d3f46eff5438ead5170f8a92604b70a181db6faf9d2092c5f4JmltdHM9MTc4MDk2MzIwMA&ptn=3&ver=2&hsh=4&fclid=2870c1ec-3641-6797-2db0-d5f537c566a5&psq=United+States+Bureau+of+Reclamation+Interagency+Agreement+%22R18PG00009%22&u=a1aHR0cHM6Ly9yZXBvc2l0b3J5LmxpYnJhcnkubm9hYS5nb3Yvdmlldy9ub2FhLzY4MTY1L25vYWFfNjgxNjVfRFMxLnBkZg)). For details on filter criteria and alternative uses of `LotekFilter` functions, see the vignette 'Advanced Usage'.


